Synthesizing Verified Components for Cyber Assured Systems Engineering \thanks{DISTRIBUTION STATEMENT A. Approved for public release.}

### 2022-09-15 paper/abstract.tex
Add notes from Junaid

Cyber-physical systems, such as avionics, must be tolerant to
cyber-attacks in the same way they are tolerant to random faults: gracefully recover or safely shut down as requirements dictate.
\brfcs\ is a model-based systems engineering environment that extends the Open Source AADL Tool Environment to add new design, analysis, and code generation capabilities for building cyber-resilient systems.
This manuscript describes a mechanism implemented by \brfcs\ to create, and insert, high-assurance components into a system for cyber-resiliency.
Such components can be filters that guard against malformed input or monitors that guard against spoofing and other malicious behavior.
A formal specification in the form of a code contract defines each high-assurance component and is used with test contracts to verify the specified behavior.
Once tested, the code contract is used to verify that the added component addresses system-level cyber requirements.
Implementations for these
high-assurance components are directly synthesized from their code
contracts and are backed up by proofs showing that high-level
specifications map in a meaning-preserving way to code.
The manuscript further reports on a case study that used \brfcs\ to cyber-harden a UAV system by inserting high assurance components to isolate the open source Air Force Research Laboratory's OpenUxAS services for
route planning.  The case study
demonstrates that synthesizing implementations from code contracts is feasible in real-world systems
engineering.

>>>>

Cyber-physical systems, such as avionics, must be tolerant to
cyber-attacks in the same way they are tolerant to random faults: gracefully recover or safely shut down as requirements dictate.
\brfcs\ is a model-based systems engineering environment that extends the Open Source AADL Tool Environment to add new design, analysis, and code generation capabilities for building cyber-resilient systems.
This manuscript describes a mechanism implemented by \brfcs\ to create, and insert, high-assurance components into a system for cyber-resiliency.
Such components can be filters that guard against malformed input or monitors that guard against spoofing and other malicious behavior.
A formal specification in the form of a code contract defines each high-assurance component and is used with test contracts for testing the specified behavior.
Once tested, the code contract is used to verify that the added component addresses system-level cyber requirements.
Implementations for these
high-assurance components are directly synthesized from their code
contracts and are backed up by proofs showing that high-level
specifications map in a meaning-preserving way to code.
The manuscript further reports on a case study that used \brfcs\ to cyber-harden a UAV system by inserting high assurance components to isolate the open source Air Force Research Laboratory's OpenUxAS services for
route planning.  The case study
demonstrates that synthesizing implementations from code contracts is feasible in real-world systems
engineering.

### 2022-09-15 paper/intro.tex
Add notes from Junaid

In recent years, aerospace stakeholders have realized that avionics systems are subject to possible cyber-attacks just like other cyber-physical systems.  
Thus, in addition to being fault-tolerant, safety-critical avionics systems must also be {\em cyber-resilient}.
Cyber-resiliency means that the system is tolerant to cyber-attacks just as safety-critical systems are tolerant to random faults: they recover and continue to execute their mission function, or safely shut down, as requirements dictate.

@@

The DARPA Cyber Assured Systems Engineering (CASE) project is targeted at developing tools for design, analysis, and verification that enable systems engineers to {\em design-in} cyber-resiliency for complex cyber-physical systems.\footnote{This work was funded in part by the Defense Advanced Research Projects Agency (DARPA).  The views expressed are those of the authors and do not reflect the official policy or position of DARPA or the U.S. Government.}
One such tool is the Model-Based Systems Engineering (MBSE) environment called {\em BriefCASE} which is based on the Architecture Analysis and Design Language (AADL)~\cite{aadl}.  BriefCASE extends the Open Source AADL Tool Environment (OSATE) to add new design, analysis, and code generation capabilities for building cyber-resilient systems.

This manuscript details how \brfcs\ facilitates inserting, specifying, testing, and synthesizing high assurance components into a system to improve its cyber-resiliency.
These are referred to as \emph{transforms} by \brfcs.
Two cyber-resiliency transformations are discussed: (1) the insertion
of a filter to prevent malformed data from a malicious actor from
being propagated to downstream components, and (2) the insertion of a
monitor to detect (and alert) unexpected behaviors arising from
untrusted components.
\brfcs\ automates inserting the components into the system, and the designer writes a \emph{code contract} for each component to define its intended behavior.

A code contract is a formal specification in the Assume Guarantee Reasoning
Environment (\agr) language.
\agr is a compositional reasoning verification engine that uses \emph{contracts} on components to specify input and output properties and then prove whether or not those properties hold when given a sub-component implementation \cite{agree2013}.
The code contract language is Turing complete allowing the designer to specify arbitrarily complex behavior.
These contracts are unit tested for correctness with \emph{test contracts}.
Test contracts define test scenarios to be implemented by the code contract under test. 
\agr\ proves whether or not that code contract implementation is correct enabling to designer to iteratively test the component behavior inside the \brfcs\ environment.
Once the behavior of the code contract is verified, \agr proves that---due to the newly
included high-assurance components---the hardened system meets its
cyber-resiliency requirements.

@@

The \brfcs\ is open source and publicly available \cite{fmide} as are the examples and case study discussed in this manuscript \cite{repo, phase2, camkes, case}.
Our approach currently applies to common cyber-vulnerabilities, such
as overflow, lack of input validation, and supply-chain issues; however, other cyber-vulnerabilities such as
side-channel attacks and denial of service are not yet dealt with in
our work.  Here, we do not report on the invention of a new type of
high-assurance component in terms of capability; instead, our contribution
is in the automated synthesis of security-improving components from
formal specifications and a means to show that the synthesis is
correct.

The approach is illustrated by a simple example in
\secref{sec:example}. \agr\ verification and specification are
presented in \secref{sec:agree} and \secref{agree-semantics}.
Code contracts are defined in \secref{sec:code-contracts} with a method
for testing code contracts in the \agr\ framework before synthesis given
\secref{sec:testing}.
The synthesis pathway is covered in Section~\ref{sec:synthesis}.

>>>>

\egm{Add to introduction and related work, that what is done at Galois called Copilot. Its a stream processing langague that generates code from the specification. Copilot to C. Effectively CodeGen from Lustre. Bounds how far needed to look in past for any value.}
In recent years, aerospace stakeholders have realized that avionics systems are subject to possible cyber-attacks just like other cyber-physical systems.  
Thus, in addition to being fault-tolerant, safety-critical avionics systems must also be {\em cyber-resilient}.
Cyber-resiliency means that the system is tolerant to cyber-attacks just as safety-critical systems are tolerant to random faults: they recover and continue to execute their mission function, or safely shut down, as requirements dictate.

@@

\egm{Remove DARPA junk.}
The DARPA Cyber Assured Systems Engineering (CASE) project is targeted at developing tools for design, analysis, and verification that enable systems engineers to {\em design-in} cyber-resiliency for complex cyber-physical systems.\footnote{This work was funded in part by the Defense Advanced Research Projects Agency (DARPA).  The views expressed are those of the authors and do not reflect the official policy or position of DARPA or the U.S. Government.}
One such tool is the Model-Based Systems Engineering (MBSE) environment called {\em BriefCASE} which is based on the Architecture Analysis and Design Language (AADL)~\cite{aadl}.  BriefCASE extends the Open Source AADL Tool Environment (OSATE) to add new design, analysis, and code generation capabilities for building cyber-resilient systems.

This manuscript details how \brfcs\ facilitates inserting, specifying, testing, and synthesizing high assurance components into a system to improve its cyber-resiliency.
These \egm{make clear what these refers to} are referred to as \emph{transforms} by \brfcs.
Two cyber-resiliency transformations are discussed: (1) the insertion
of a filter to prevent malformed data from a malicious actor from
being propagated to downstream components, and (2) the insertion of a
monitor to detect (and alert) unexpected behaviors arising from
untrusted components.
\brfcs\ automates inserting the components into the system, and the designer writes a \emph{code contract} for each component to define its intended behavior.

A code contract is a formal specification in the Assume Guarantee Reasoning
Environment (\agr) language.
\agr\ is a compositional reasoning verification engine that uses \emph{contracts} on components to specify input and output properties and then prove whether or not those properties hold when given a sub-component implementation \cite{agree2013}.
The code contract language is Turing complete allowing the designer to specify arbitrarily complex behavior.
These contracts are unit tested for correctness with \emph{test contracts}.
Test contracts define test scenarios to be implemented by the code contract under test. 
\agr\ proves whether or not that code contract implementation is correct, enabling the designer to iteratively test the component behavior inside the \brfcs\ environment.
Once the behavior of the code contract is verified, \agr\ proves that---due to the newly
included high-assurance components---the hardened system meets its
cyber-resiliency requirements.

@@

\brfcs\ is open source and publicly available \cite{fmide} as are the examples and case study discussed in this manuscript \cite{repo, phase2, camkes, case}.
Our approach currently applies to common cyber-vulnerabilities, such
as overflow, lack of input validation, and supply-chain issues; however, other cyber-vulnerabilities such as
side-channel attacks and denial of service are not yet dealt with in
our work.  Here, we do not report on the invention of a new type of
high-assurance component in terms of capability; instead, our contribution
is in the automated synthesis of security-improving components from
formal specifications and a means to show that the synthesis is
correct.

We now give a summary of the contributions detailed in this manuscript.
\begin{compactitem}
  \item The language and semantics of code contracts to specify the behavior of high assurance components.
  \item Test contracts as a stylized mechanism to unit test code contracts inside the \brfcs\ framework with \agr.
  \item A synthesis path from code contracts to \ckml\ with a formal framework to argue correctness.
  \item A formal argument that \agr\ verification of code contracts carry over to the resulting binaries from the \ckml\ compiler.
  \item A case study that used the implementation of the approach in \brfcs\ to add cyber-resiliency to a non-trivial UAV system.
\end{compactitem}
The rest of this manuscript is organized as follows.
\secref{sec:overview} is an overview of the \brfcs\ environment and key tools in that environment relative to the contributions of this work. The approach is illustrated by a simple example in
\secref{sec:example}. Essential background on \agr\ verification and specification are
presented in \secref{sec:agree}.
The language and semantics of code contracts are defined in \secref{sec:code-contracts}, and test contracts are defined in \secref{sec:testing}.
The synthesis pathway is covered in Section~\ref{sec:synthesis}. 
\secref{sec:case-study} discusses the case study.
This is followed by related work in \secref{sec:related-work}. 
The conclusions and future work are in \secref{sec:conclusion}.
